---
title: "Rmarkdown test"
author: "EK Lipp"
date: "May 16, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# MERS Analysis

This a report of the visualization examples prepared as part of IDEAS workshop day 1. 

```{r mers}
library(lubridate)
library(ggplot2)
library(plotly)

mers <- read.csv('cases.csv')
class(mers$onset)
mers$hospitalized[890] <-('2015-02-20')
mers <- mers[-471,]
mers$onset2 <- ymd(mers$onset)
mers$hospitalized2 <- ymd(mers$hospitalized)
class(mers$onset2)
day0 <- min(na.omit(mers$onset2))
mers$epi.day <- as.numeric(mers$onset2 - day0)
```

## Global analysis of MERS



```{r, echo=FALSE}
ggplot(data=mers) + 
  geom_bar(mapping=aes(x=epi.day)) +
  labs(x='Epidemic day', y='Case count', 
       title='Global count of mERS cases by date of symptom onset', 
       caption="Data from https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```


The following chart allows us to examine cases by country

```{r, echo=FALSE}
ggplot(data=mers) + 
  geom_bar(mapping=aes(x=epi.day, fill=country)) +
  labs(x='Epidemic day', y='Case count', 
       title='Global count of mERS cases by date of symptom onset', 
       caption="Data from https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```

## MERS infectious period

### Univariate analyses

```{r, echo=FALSE}
mers$infectious.period <- mers$hospitalized2-mers$onset2
class(mers$infectious.period)
mers$infectious.period <- as.numeric(mers$infectious.period, units="days")

ggplot(data=mers) + 
  geom_histogram(aes(x=infectious.period)) +
  labs(x='Infectious Period', y='Frequnecy',title='Distribution of calculated MERS infectious period', 
       caption="Data from: https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```
The above example incluced nosocomial infections. Below we remove any cases that start before hospitalization


```{r, echo=FALSE}
mers$infectious.period2 <- ifelse(mers$infectious.period<0,0,mers$infectious.period)

ggplot(data=mers) +
  geom_histogram(aes(x=infectious.period2)) +
  labs(x='Infectious period', y='Frequency',
       title='Distribution of calcuated MERS infectious period (pos values only)',caption="Data from https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```

### Bivariate analyses

The following replicates the orginal graphs but with nosocomial infections removed.

```{r, echo=FALSE}
ggplot(data=mers) + 
  geom_point(mapping=aes(x=epi.day, y=infectious.period2, colour=country)) +
  labs(x='Epidemic day', y='Case count', 
       title='Global count of mERS cases by date of symptom onset', 
       caption="Data from https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```

The following explores data using facets (broken out my country)

```{r, echo=FALSE}
ggplot(data=mers, mapping=aes(x=epi.day, y=infectious.period2)) +
  geom_point(mapping = aes(color=country)) +
  facet_wrap(~ country) +
  scale_y_continuous(limits = c(0,50)) +
  labs(x='Epidemic day',y="Infectious period", 
       title='MERS infectious period (pos values ony) over time', 
       caption="Data from https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
```

##Interactive chart with plotly


```{r, echo=FALSE}
f<- list(
  family="Courier New, monospace",
  size = 18,
  color="#7f7f7f"
)
x <- list(
  title = "Infectious Period",
  titlefont = f
)
y <- list(
  title = "Epidemic Day",
  titlefont = f
)
p<- plot_ly(x = mers$infectious.period2, y = mers$epi.day, type='histogram') %>% layout(xaxis = x, yaxis = y)
p
```
### Interactive chart using ggplotly

```{r, echo=FALSE}
onset<- ggplot(data=mers) + 
  geom_bar(mapping=aes(x=epi.day, fill=country)) +
  labs(x='Epidemic day', y='Case count', 
       title='Global count of mERS cases by date of symptom onset', 
       caption="Data from https://github.com/rambaut/MERS-Cases/blob/gh-pages/data/cases.csv")
ggplotly(onset)
```
